---
title: "Prediction Assignment Writeup"
author: "Kowshik Sarker"
date: "Jul 13, 2020"
output: html_document
---

The goal of the project was to fit a Machine Learning model in order to predict the way some weight exercises were performed given data about the movement of subjects. For full information about the data, please visit http://groupware.les.inf.puc-rio.br/har

# Understanding the Data
We first load the dataset and understanting about the data.
```{r Load Libraries and Data, warning=FALSE, message=FALSE, cache=FALSE}
library(caret)
library(dplyr)
library(RCurl)
library(gridExtra)
library(corrplot)
training <- read.csv("data/pml-training.csv")
testing <- read.csv("data/pml-testing.csv")
dim(training)
dim(testing)
```

## Choosing Predictor Variables
Because of the number of variables (160, the model was taking too long to build and so we decided to use the data that had complete data.  These included belt, arm, dumbbell and forearms)

For more information about the dataset go here >> http://groupware.les.inf.puc-rio.br/work.jsf?p1=11201

Here are the predictor variables selected

```{r GoodDataSearch}
removeMissingData <- sapply(testing, function (x) any(is.na(x) | x == ""))
definingPredictorVariables <- !removeMissingData & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(removeMissingData))
PredictorVariables <- names(removeMissingData)[definingPredictorVariables]
PredictorVariables
```

We have reduced the variables to analyze significantly.
```{r SubsettingNonPredictors}
onlyPredictors <- c("classe", PredictorVariables)
training <- training[, onlyPredictors]
dim(training)
```


# Build the model
The model uses k-mode cross validation and is run 3 times on 10-fold data starting with 70% of the data for training and 30% for validating the model.

Below the accuracy will test the prediction accuracy derived from the training data against the validation data
```{r, cache=TRUE, warning=FALSE}
nuFoldData <- 10
nuRepeats <- 3
testDataSize <- 0.7
base <- read.csv("data/pml-training.csv")
tempTraining <- createDataPartition(base$classe, p = testDataSize, list = FALSE)
training <- base[tempTraining,]
validation <- base[-tempTraining,]
#Using all variables
tempTrControl <- trainControl(method = "repeatedcv", number = nuFoldData, repeats = nuRepeats)
trainedModel <- train(classe ~ ., data = training[,c("classe", "roll_belt","roll_arm", "roll_dumbbell", "roll_forearm", "pitch_belt", "pitch_arm", "pitch_dumbbell", "pitch_forearm", "yaw_belt", "yaw_arm", "yaw_dumbbell", "yaw_forearm")],
                method = "rf", ntree = nuFoldData, trControl = tempTrControl)
confusionMatrix(trainedModel, newdata = predict(trainedModel, newdata = validation))
```

## Predicting events using testing data
Let's test our data using the test data given above
```{r TestingTheModel, warning=FALSE, message=FALSE}
nuPredictions <- 20
finalData <- rep(NA, nuPredictions)
for(i in 1:nuPredictions){
  tmpID <- filter(testing, problem_id == i)
  finalData[i] <- as.vector(predict(trainedModel, newdata = tmpID))
}
finalData
```